Kindly Bent to Free Us

### 2019-11-18 kindly/motivation.tex
Misc compression.

\lang{} is amenable to a functional resource-passing style as it is
common in functional encodings of session types (e.g.,
\cite{DBLP:journals/jfp/Padovani17}; see also
\cref{sec:session-linearity} in the supplement). But it really shines 
when manipulating mutable resources like buffers or connection pools
using a mixed functional and imperative programming style.
%
To support this usage pattern of linearity, we rely on
the notion of borrowing \cite{DBLP:conf/popl/BoylandR05} which
we demonstrate through a series of examples:
\begin{itemize}
\item An API for linear arrays to demonstrate simple borrowing and imperative programming.
\item
  A backtracking Sudoku solver to demonstrate
  \lang's ability to build safe hybrid functional/imperative data-structures.
\item A pool of linear objects to demonstrate advanced usage of regions.
\end{itemize}

@@

\begin{figure}[tp]
  \centering
  \begin{subfigure}[t]{1\linewidth}
\begin{lstlisting}
type ('a:'k) pool : 'k

create : (unit -> 'a) -> 'a pool
consume : ('a -> unit) -> 'a pool -> unit
use : ('a_1:'k_1),('a_2:'k_2),('k_2 <= aff_r) =>
  &('a_1 pool) -> (&!(aff_r1,'a_1) -{lin}> 'a_2) -{'k_1}> 'a_2
\end{lstlisting}
    \vspace{-10pt}
    \caption{Signature}
    \label{intf:pool}

@@

  \begin{subfigure}[t]{1\linewidth}
\begin{lstlisting}
type ('a:'k) pool : 'k = {
  spawn : unit -> 'a ;
  queue : 'a CQueue.t;
}

let create spawn =
  { spawn ; queue = CQueue.create () }
let consume f c =
  CQueue.iter f c.queue

let use { spawn ; queue } f =
  let o = match CQueue.pop &queue with
    | Some x -> x
    | None () -> spawn ()
  in
  let r = {| f &!o |} in(*@\label{line:pool:region}*)
  Queue.push o &queue;
  r
\end{lstlisting}
    \vspace{-15pt}
    \caption{Implementation}
    \label{impl:pool}
  \end{subfigure}

>>>>

\lang{} is amenable to a functional resource-passing style
common in functional encodings of session types (e.g.,
\cite{DBLP:journals/jfp/Padovani17}; see also
\cref{sec:session-linearity} in the supplement). But it really shines 
when manipulating mutable resources like buffers or connection pools
using a mixed functional and imperative programming style.
%
To support this usage pattern of linearity, we rely on
the notion of borrowing \cite{DBLP:conf/popl/BoylandR05} which
we demonstrate through a series of examples:
\begin{itemize}
\item An API for linear arrays to demonstrate simple borrowing and imperative programming.
\item
  A backtracking Sudoku solver to demonstrate
  \lang's ability to build safe hybrid functional/imperative data-structures.
\item A pool of linear objects to demonstrate advanced use of regions.
\end{itemize}

@@

\begin{figure}[tp]
  \centering
  \begin{subfigure}[t]{1\linewidth}
\begin{lstlisting}
type ('a:'k) pool : 'k
create : (unit -> 'a) -> 'a pool
consume : ('a -> unit) -> 'a pool -> unit
use : ('a_1:'k_1),('a_2:'k_2),('k_2 <= aff_r) =>
  &('a_1 pool) -> (&!(aff_r1,'a_1) -{lin}> 'a_2) -{'k_1}> 'a_2
\end{lstlisting}
    \vspace{-10pt}
    \caption{Signature}
    \label{intf:pool}

@@

  \begin{subfigure}[t]{1\linewidth}
\begin{lstlisting}
type ('a:'k) pool : 'k = {spawn: unit -> 'a; queue: 'a CQueue.t}
let create spawn = { spawn ; queue = CQueue.create () }
let consume f c = CQueue.iter f c.queue
let use { spawn ; queue } f =
  let o = match CQueue.pop &queue with
    | Some x -> x
    | None () -> spawn ()
  in
  let r = {| f &!o |} in(*@\label{line:pool:region}*)
  Queue.push o &queue;
  r
\end{lstlisting}
    \vspace{-15pt}
    \caption{Implementation}
    \label{impl:pool}
  \end{subfigure}
